\documentclass[a4]{article}
\usepackage[total={6.7in,10.2in}, % text area width and height
		    top=1in, left=1in, right=1in, bottom=1in,% top and left margin
		    includefoot]{geometry}
\usepackage{rsllisting}
\lstset{language=rsl}

\title{Course 02263 Mandatory Assignment 2, 2018} 

\author{Marc Storm Larsen (s144452),\\ 
        Mario Padilla Carrion (s172387)}

\begin{document}

\maketitle

\tableofcontents
\newpage

\section{Work distribution}
The work carried out during the analysis, implementation and testing of the program and this paper, has been distributed equally between the both of us, where each part of the work has been discussed with the other afterwards. What is meant by that is when one had carried out for example an analysis of part of the program, it was discussed with the other member, such that the both of us had their opinions expressed as well as ensuring that the analysis was comprehensive. This way of working has been used for all parts of the project, to ensure that both of us had part in every part of the project and to make sure that we both understand the entire solution.

\section{Introduction}
This document contains a solution to the problem of representing trams nets and tram time tables, where the trams in a net must follow a plan. These nets and time tables will be represented with new data types. Furthermore will the document describe a number of functions, that can be used to check whether concrete values of the data types to describe nets and time tables satisfies a number of requirements, making them wellformed. Lastly will the document describe the test cases that were carried out, to prove the correctness and robustness of the solution.

\section{Formal Specification of Nets}

\lstinputlisting{assignment-2/src/NET.rsl}

\noindent To model a \emph{net}, a few types have been added, one of which is \verb=Connection=, which collects information about a connection in one direction in a net. The information it collects is what stops it is connecting, the headway, the maximum capacity and the minimum time a tram should use to drive on the connection. Another is \verb=Connections= which is a type used to associate a \emph{stop} with a set of \emph{connections}, a \verb=set= is used as the order of the connections is irrelevant. Another is \emph{Stops} which is used to associate a \emph{stop} with its \emph{capacity}. Lastly is the type \verb=Net= used to collect the \emph{stops} and \emph{connections} in a \emph{net}. Below are the requirements for a \emph{net} defined.
\\
\begin{description}
    \item[isWellformed] To the say that a \emph{net} is wellformed, means that is should fulfil the following requirements. Each requirement will indicate what function is used to ensure the requirement.
        \begin{itemize}
            \item The \emph{net} must contain at least two \emph{stops}. Function: \verb=hasCorrectSize=.
            \item Any \emph{stop} must have a direct or indirect connection to every other \emph{stop} in the \emph{net}, i.e. a tram can travel to any \emph{stop} from any \emph{stop}. Function: \verb=isNetConnected=.
            \item A \emph{stop} can not have a direct \emph{connection} to itself. Function: \verb=noDirectSelfConnections=.
            \item A \emph{stop} can not have more than one direct \emph{connection} to the same \emph{stop}. Function: \verb=hasOneDirectConnection=.
            \item A \emph{connection} between two \emph{stops} consists of two tracks, one track for each direction, i.e. a tram can travel back and forth between the two \emph{stops} using the same \emph{connection}. Function: \verb=hasBiDirectConnection=.
        \end{itemize}
\end{description}

\noindent In the list below one will find a description of all the functions. If one of the functions utilises the functionality of another function, it will be stated in the end of the description. \\

\begin{description}
    \item[hasCorrectSize] \hfill \\ Can be used to check whether or not a \emph{net} has a correct size, i.e. the number of \emph{stops} in the \emph{net}.
    \item[isNetConnected] Can be used to check that the \emph{net} is connected, i.e. there is a direct or indirect \emph{connection} from any \emph{stop} to all other \emph{stops}. Function: \verb=areConnected=.
    \item[noDirectSelfConnections] \hfill \\ Can be used to check that any \emph{stop} in the \emph{net} does not have a direct \emph{connection} to itself. Function: \verb=areDirectlyConnected=.
    \item[hasOneDirectConnection] \hfill \\ Can be used to check if any \emph{stop} in the \emph{net} only has one direct \emph{connection} to another \emph{stop}. Function:  \verb=Connections=.
    \item[hasBiDirectConnection] \hfill \\ Can be used to check that all direct \emph{connections} between two \emph{stops} in the \emph{net} are connecting the \emph{stops} in both directions. Function: \verb=areDirectlyConnected=.
    \item[isIn] \hfill \\ Can be used to check whether or not a \emph{stop} is in a \emph{net}.
    \item[capacity - stop] \hfill \\ Can be used to find out what the capacity of a \emph{stop} is.
    \item[areDirectlyConnected] \hfill \\ Can be used to check whether or not two \emph{stops} are directly connected in a \emph{net}. Function: \verb=hasDirectConnection=.
    \item[minHeadway] \hfill \\ Can be used to find out what the minimum \emph{headway} between two connected \emph{stops} is. Function: \verb=getConnection=.
    \item[capacity - connection] \hfill \\ Can be used to find out what the capacity of a \emph{connection} is. Function: \verb=getConnection=.
    \item[minDrivingTime] \hfill \\ Can be used to find out what the minimum driving time between two connected \emph{stops} is. Function: \verb=getConnection=.
    % Our own observer functions
    \item[getConnection] \hfill \\ Can be used to get the \emph{connection} between two directly connected \emph{stops}. Function: \verb=filterConnections=.
    \item[filterConnections] \hfill \\ Can be used to find a single \emph{connection} in a set of connections.
    \item[hasDirectConnection] \hfill \\ Can be used to check if two \emph{stops} are directly connected in one direction in a \emph{net}.
    \item[areConnected] Can be used to check if two \emph{stops} are connected, either with a direct or indirect \emph{connection}.
\end{description}  

\section{Formal Specification of Time Tables}

\lstinputlisting{assignment-2/src/TIMETABLE.rsl}

\noindent It is assumed that two consecutive rows in a \emph{plan} for a \emph{tram} is implying that a tram is driving from stop $x$ to $y$ without passing any intermediate stops, i.e. driving to a stop via a direct connection. Furthermore should the times in any \emph{plan} in a \emph{time table} be in increasing order. This means that should a \emph{tram} drive from \emph{stop} $x$ at 1 hour and 15 min, it should be written as 75 min. \\

The type $TimeTable$ is created to associate a \verb=Tram= with a verb=Plan=, where a \verb=Plan= is a list of the type \verb=TimeSlot=. A \verb=TimeSlot= contains information about what \emph{stop} the tram should be at, as well as what time it should arrive and depart at the given \emph{stop}.
From the \verb=TimeTable= that is created by the user and is wished to be checked whether it is wellformed or not are two data structures created. These data structures extract information from the \emph{plans} in the \emph{time table} to present the data in a nice and more manageable way for the functions that have been created to check the wellformness. The general idea of the two data structures \verb=PlanConns= and \verb=PlanStops= is to store the arrival and departure times in separate lists for connections and stops respectively. E.g. this means that all the arrival and departure times to the connection between stop $A$ and stop $B$ can be easily found, as well as all the arrival and departure times at stop $A$. \verb=PlanSlots= is used to store these arrival and departure times, where a list is used to keep the ordering of the times. \verb=PlanStops= is used to associate a \emph{stop} with a \verb=PlanSlots= and \verb=PlanConns= is used to associate a connection with a \verb=PlanSlots=. The ID of a connection is derived from the two stops it is connecting, where the direction matter. So a connection between stop $x$ and $y$ will have ID $xy$ for the track on the connection going from $x$ to $y$ and vice versa. The times in the \verb=PlanSlots= are derived from the time a tram arrives and departs from a \emph{stop}. Say a \emph{tram} departs from \emph{stop} $x$ at time $a$ and arrives at \emph{stop} $y$ at time $b$, then we say the \emph{tram} arrives at connection $xy$ at time $a$ and departs at time $b$. \\

The main reason not to create the \verb=TimeTable= directly with these two types, was to reduce the amount of updates to the two types and instead create them once the \verb=TimeTable= has been finalised by the user and wants to check whether it is wellformed.
\\

\begin{description}
    \item[isWellformed] To the say that a \emph{Time Table} is wellformed, means that is should fulfil the following requirements. Each requirement will indicate what function is used to ensure the requirement.
        \begin{itemize}
            \item All times in a \emph{time table} must be equal or increasing, i.e. a tram can not travel back in time. Function: \verb=hasCorrectTimes=
            \item All stops that is a part of a \emph{time table} must be part of the \emph{net}. Function: \verb=allStopsInNet=
            \item Any two consecutive rows in a \emph{time table} implies that the \emph{tram} is driving directly between the \emph{stops} on a direct connection, the direct connection should therefore exists in the \emph{net}. Function: \verb=noIndirectConnections=
            \item The capacity of any \emph{stop} in the \emph{net} must not be exceeded. Function: \verb=noStopCapacityExceeded=
            \item The capacity of any \emph{connection} in either direction in the \emph{net} must not be exceeded. Function: \verb=noConnectionCapacityExceeded=
            \item The time between two \emph{trams} arriving and departing from a \emph{stop} must not be shorter then the headway of the connection they were arriving and departing from. Function: \verb=noHeadwayViolation=
            \item Any \emph{tram} is not allowed to drive faster on a connection between two \emph{stops} than the connection allows. Function: \verb=noDrivingTimeExceeded=
        \end{itemize}
\end{description}

\noindent In the list below one will find a description of all the functions. If one of the functions utilises the functionality of another function, it will be stated in the end of the description. \\

\begin{description}
    \item[hasCorrectTimes - Time Slot list] \hfill \\ This function can be used to check that all times in a timetable are equal or increasing. 
    \item[hasCorrectTimes - Time Table] \hfill \\ This function can be used to check that all times in \emph{time tables} for all \emph{trams} are correct. Function: \verb=hasCorrectTimes - Time Slot list=
    \item[allStopsInNet] \hfill \\ This function can be used to check that all stops in a \emph{time table} is part of the \emph{net}. Function: \verb=isIn - Net=
    \item[noIndirectConnections] \hfill \\ This function can be used to check that there exists a direct connection between two stops in the \emph{net} for any two consecutive rows in a \emph{time table}. Function: \verb=hasDirectConnection=
    \item[noStopCapacityExceeded] \hfill \\ This function can be used to check that a \emph{time table} does not exceed the capacity of any \emph{stop} in a \emph{n}. Functions: \verb=findMaxCapacity=, \verb=capacity - stop=
    \item[noConnectionCapacityExceeded] \hfill \\ This function can be used to check that a \emph{time table} does not exceed the capacity of any \emph{connection} in a \emph{n}. Functions: \verb=findMaxCapacity=, \verb=capacity - connection=
    \item[noHeadwayViolation - list] \hfill \\ This function can be used to check that trams are not arriving at or departuring from a \emph{stop} too close to each other. Function: \verb=minHeadway=
    \item[noHeadwayViolation] \hfill \\ This function can be used to check that a \emph{time table} does not having \emph{trams} arrive from and depart at \emph{stops} too close to each other. Function: \verb=noHeadwayViolation - list=
    \item[noDrivingTimeExceeded] \hfill \\ This function can be used to check that a \emph{time table} does not have any \emph{trams} drive faster between any \emph{stops} than what is allowed. Function: \verb=minDrivingTime=
\end{description}

% ------------ GENERATORS ------------
% Describe the general idea of the two need datastructure.
\begin{description}
    \item[createPlanConns] \hfill \\ Creates the datastructure of the plan of the \emph{connections} from a \emph{time table}. Function: \verb=createPlanConnsRec=
    \item[createPlanConnsRec] \hfill \\ Creates the datastructure of the plan of the \emph{connections} from all the \emph{trams} from a \emph{time table}. Function: \verb=createPlanConnsRec=
    \item[createPlanConnsForPlan] \hfill \\ Creates the plan of the \emph{connections} for a single \emph{tram}. Functions: \verb=createPlanConnsForPlan=, \\ \verb=insertPlanSlot - connection=
    \item[insertPlanSlot - connection] \hfill \\ Inserts the plan in the datastructure. 
    \item[createPlanStops] \hfill \\ Creates the datastructure of the plan of the \emph{stops} from a \emph{time table}. Function: \verb=createPlanStopsRec=
    \item[createPlanStopsRec] \hfill \\ Creates the datastructure of the plan of the \emph{stops} from all the \emph{trams} from a \emph{time table}. Function: \verb=createPlanStopsRec=
    \item[createPlanStopsForPlan] \hfill \\ Creates the plan of the \emph{stops} for a single \emph{tram}. Functions: \verb=createPlanStopsForPlan=, \\ \verb=insertPlanSlot - stop=
    \item[insertPlanSlot - stop] \hfill \\ Inserts the plan in the datastructure.
\end{description}

% ------------ OBSERVERS ------------
\begin{description}
    \item[isIn] \hfill \\ Can be used to check whether a \emph{tram} exists in a \emph{time table}
    \item[isInPlan - Stop] \hfill \\ Can be used to check whether a \emph{stop} exists in a plan.
    \item[isInPlan - Connection] \hfill \\ Can be used to check whether a \emph{connection} exists in a plan
    \item[findMaxCapacity] \hfill \\ Can be used to find the max capacity that is used for two time ranges. Function: \verb=findMaxCapacityRec=
    \item[findMaxCapacityRec] \hfill \\ Can be used to run through two time ranges and calculate the maximum capacity that is needed for those, i.e. the maximum overlap. Function: \verb=findMaxCapacityRec=
    \item[sort] \hfill \\ Can be used to sort a list of times. Functions: \verb=sort=, \verb=getSmallerItems=, \verb=getLargerItems=
    \item[getLargerItems] \hfill \\ Can be used the find times that are larger then a given time. Function: \verb=getLargerItems=
    \item[getSmallerItems] \hfill \\ Can be used the find times that are smaller or equal then a given time. Function: \verb=getSmallerItems=
    \item[isSorted] \hfill \\ Can be used to check whether a list of times is sorted in increasing order
\end{description}

\section{Testing the Specifications}

\subsection{Nets}

\lstinputlisting{assignment-2/src/NETTest.rsl}

\subsection{Time Tables}

\lstinputlisting{assignment-2/src/TIMETABLETest.rsl}

\section{Testing by translation to SML}

\subsection{Test specification}

The purpose of each individual test has been added as source code comments and can be seen in the contents of the \verb=NETTest.rsl=  and \verb=TIMETABLETest.rsl= files, which are shown above. The tests have been divided into three sections, corresponding to \emph{Generators, Observers and Predicates} with their respective auxiliary functions, to increase the readability.\\

The general idea of the approach to test the "atomic"/individual functions, have been as follows. For the functions that returns a \verb=Bool= value, we have created test data for the test cases, where we expect the result to be true. Furthermore, did we have a lot of focus on creating invalid test data, to test that the functions were correctly returning false, to test their robustness. As for functions that returns other values than \verb=Bool=. 
Here has there also been created test data for test cases, where the result has been compared to the expected output, to get a truth value in the overview of the test results, which can be seen in the following section.\\

The test data created for these functions have both been best case test data, where everything is as it should be, but also some of the test data has not been best case, to for example test: wrong Nets generated with for example \emph{Connections} between \emph{Stops} that are not part of the \emph{Net} and to test also, wrong \emph{TimeTables} generated with for example \emph{Plans} with \emph{arrival} time before \emph{departure} time, i.e., travelling back in time.

\subsection{Results of evaluations}

The results of evaluating the SML translation of the RSL test cases for both the \verb=NET.rsl= and \verb=TIMETABLE.rsl= are that all tests are returning the expected values and thereby every tests run successfully. For a complete overview of the result of the individual test see appendix~\ref{app:results}.

From the results of the tests and the implementation itself we can conclude that we have successfully created an application that can be used to check the wellformedness of a trams net and a tram time table, where trams in the net must follow a plan. The application can successfully check that no trams drive to fast between two stops, leaves enough headway to each other and do not exceed the capacities of both the connections and stops.


\appendix
\section{Test results} \label{app:results}
\begin{table}[h]
    \begin{tabular}{ll}
        \multicolumn{2}{l}{NETTest.rsl}          \\
        \multicolumn{1}{l|}{Test case}  & Result \\ \hline 
        \multicolumn{1}{l|}{{[}t100{]}} & true   \\
        \multicolumn{1}{l|}{{[}t101{]}} & true   \\
        \multicolumn{1}{l|}{{[}t200{]}} & true   \\
        \multicolumn{1}{l|}{{[}t201{]}} & true   \\
        \multicolumn{1}{l|}{{[}t202{]}} & true   \\
        \multicolumn{1}{l|}{{[}t300{]}} & true   \\
        \multicolumn{1}{l|}{{[}t301{]}} & true   \\
        \multicolumn{1}{l|}{{[}t302{]}} & true   \\
        \multicolumn{1}{l|}{{[}t400{]}} & true   \\
        \multicolumn{1}{l|}{{[}t500{]}} & true   \\
        \multicolumn{1}{l|}{{[}t600{]}} & true   \\
        \multicolumn{1}{l|}{{[}t601{]}} & true   \\
        \multicolumn{1}{l|}{{[}t700{]}} & true   \\
        \multicolumn{1}{l|}{{[}t800{]}} & true   \\
        \multicolumn{1}{l|}{{[}t900{]}} & true   \\
        \multicolumn{1}{l|}{{[}t1000{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1100{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1200{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1300{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1301{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1400{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1401{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1402{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1403{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1500{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1501{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1502{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1503{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1600{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1601{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1602{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1700{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1800{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1801{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1900{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1901{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1902{]}} & true  
    \end{tabular}
\end{table}

\begin{table}[h]
    \begin{tabular}{ll}
        \multicolumn{2}{l}{TIMETABLETest.rsl}   \\
        \multicolumn{1}{l|}{Test case}  & Result \\ \hline 
        \multicolumn{1}{l|}{{[}t100{]}} & true   \\
        \multicolumn{1}{l|}{{[}t101{]}} & true   \\
        \multicolumn{1}{l|}{{[}t200{]}} & true   \\
        \multicolumn{1}{l|}{{[}t201{]}} & true   \\
        \multicolumn{1}{l|}{{[}t202{]}} & true   \\
        \multicolumn{1}{l|}{{[}t300{]}} & true   \\
        \multicolumn{1}{l|}{{[}t301{]}} & true   \\
        \multicolumn{1}{l|}{{[}t400{]}} & true   \\
        \multicolumn{1}{l|}{{[}t401{]}} & true   \\
        \multicolumn{1}{l|}{{[}t402{]}} & true   \\
        \multicolumn{1}{l|}{{[}t500{]}} & true   \\
        \multicolumn{1}{l|}{{[}t501{]}} & true   \\
        \multicolumn{1}{l|}{{[}t502{]}} & true   \\
        \multicolumn{1}{l|}{{[}t600{]}} & true   \\
        \multicolumn{1}{l|}{{[}t601{]}} & true   \\
        \multicolumn{1}{l|}{{[}t700{]}} & true   \\
        \multicolumn{1}{l|}{{[}t701{]}} & true   \\
        \multicolumn{1}{l|}{{[}t702{]}} & true   \\
        \multicolumn{1}{l|}{{[}t800{]}} & true   \\
        \multicolumn{1}{l|}{{[}t801{]}} & true   \\
        \multicolumn{1}{l|}{{[}t802{]}} & true   \\
        \multicolumn{1}{l|}{{[}t900{]}} & true   \\
        \multicolumn{1}{l|}{{[}t901{]}} & true   \\
        \multicolumn{1}{l|}{{[}t902{]}} & true   \\
        \multicolumn{1}{l|}{{[}t903{]}} & true   \\
        \multicolumn{1}{l|}{{[}t1000{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1100{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1200{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1300{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1301{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1302{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1303{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1400{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1401{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1500{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1501{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1600{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1601{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1602{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1603{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1604{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1700{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1701{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1800{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1801{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1802{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1900{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1901{]}} & true  \\
        \multicolumn{1}{l|}{{[}t1902{]}} & true
    \end{tabular}
\end{table}

\begin{table}[h]
    \begin{tabular}{ll}
        \multicolumn{2}{l}{TIMETABLETest.rsl}   \\
        \multicolumn{1}{l|}{Test case}  & Result \\ \hline 
        \multicolumn{1}{l|}{{[}t2000{]}} & true  \\
        \multicolumn{1}{l|}{{[}t2001{]}} & true  \\
        \multicolumn{1}{l|}{{[}t2100{]}} & true  \\
        \multicolumn{1}{l|}{{[}t2101{]}} & true  \\
        \multicolumn{1}{l|}{{[}t2200{]}} & true  \\
        \multicolumn{1}{l|}{{[}t2300{]}} & true  \\
        \multicolumn{1}{l|}{{[}t2301{]}} & true  \\
        \multicolumn{1}{l|}{{[}t2302{]}} & true 
    \end{tabular}
\end{table}

\end{document}
